VM-Centric Snapshot Deduplication with Fault Isolation for Converged Cloud Architectures

### 2014-07-17 snapshot_system/paper_dedup.tex
rev

% {\it [Need to find a place to put these numbers in: Total number of chunks 
% in 350 snapshots: 1,546,635,485. 
% Total number of chunks after localized dedup: 283,121,924. Total number of unique chunks: 87,692,682.]}
Our key observation is for those chunks that appear in different VMs, top $k$ most popular items
dominate the distribution.  This popular data set is called the \textbf{PDS}. 
Figure~\ref{fig:Datazipf} shows the distribution of chunk popularity for a data trace 
from Alibaba with 2500 VMs discussed in Section ~\ref{sect:evaluation}.
We define chunk popularity as the number of unique copies of the chunk in the data-store,
i.e., the number of copies of the chunk after local deduplication.
%the distribution of chunk popularity~\cite{WeiZhangIEEE} 
%based on a production data trace from Alibaba with over 2500 VMs.
The distribution from this figure is Zipf-like. 
Let  $\sigma$ denote as the percentage of unique data chunks belonging to PDS and 
from the evaluation in Section~\ref{sect:evaluation}, we find that
$\sigma$ with a range of 2 to 4\% can deliver a fairly competitive deduplication efficiency.
%For example,
%with top 2\% of the most popular chunks stored in PDS, up-to 40\% of the raw backup 
%data can be eliminated by fingerprint lookup in PDS index.

>>>>

% {\it [Need to find a place to put these numbers in: Total number of chunks 
% in 350 snapshots: 1,546,635,485. 
% Total number of chunks after localized dedup: 283,121,924. Total number of unique chunks: 87,692,682.]}
Our key observation is for those chunks that appear in different VMs, top $k$ most popular items
dominate the distribution.  This popular data set is called the \textbf{PDS}. 
Figure~\ref{fig:Datazipf} shows the distribution of chunk popularity for a data trace 
from Alibaba with 2500 VMs discussed in Section ~\ref{sect:evaluation}.
We define chunk popularity as the number of unique copies of the chunk in the data-store,
i.e., the number of copies of the chunk after local deduplication.
%the distribution of chunk popularity~\cite{WeiZhangIEEE} 
%based on a production data trace from Alibaba with over 2500 VMs.
The distribution from this figure is Zipf-like. 
Let  $\sigma$ denote as the percentage of unique data chunks belonging to PDS and 
from the evaluation in Section~\ref{sect:evaluation}, we find that
$\sigma$ with about  2\% can deliver a fairly competitive deduplication efficiency.
%$\sigma$ with about  range of 2 to 4\% can deliver a fairly competitive deduplication efficiency.
%For example,
%with top 2\% of the most popular chunks stored in PDS, up-to 40\% of the raw backup 
%data can be eliminated by fingerprint lookup in PDS index.

### 2014-07-17 snapshot_system/paper_eval.tex
add deletion
rev

\subsection{Deduplication Efficiency}
\label{sect:evaldedup}

@@

\comments{
alg     efficiency
\scriptsize{}VC-2\%\tiny{}No\ Similarity        94.3132243792193
\scriptsize{}VC-NoPDS   93.0202119345744
\scriptsize{}VC-2\%PDS  96.0127647380588
\scriptsize{}VC-4\%PDS  96.5784837160482
\scriptsize{}SRB        97.8625610883421
}
\begin{table}[htb]
\centering
\begin{small}
    \begin{tabular}{|c|c|c|c|c|c|}
    \hline 
&
     \multicolumn{4}{|c|}{VC}  & VO \\ \hline
&
     2\%PDS  &  No PDS &2\%PDS     & 4\%PDS & SRB  \\ 
&     no-simi&   &&  &    \\ \hline
Dedup &     94.31\%& 93.02\%  & 96.01\% & 96.58\% & 97.86\% \\ \hline
Mem(MB) &     108&  46 & 118  & 190 & 2400 \\ \hline
    \end{tabular}
\end{small}
    \caption{Deduplication efficiency and per-machine memory usage for different VC settings and
        Stateless Routing with Binning (SRB).}
    \label{tab:efficiency-comparison}
\end{table}

Table~\ref{tab:efficiency-comparison} shows the deduplication efficiency for SRB and VC with different 
settings.
% <<<<<<< HEAD
% Our deduplication efficiency is defined as the percent of duplicate chunks
% which are detected and deduplicated, so only global perfect deduplication can have 100\% efficiency.
% The figure also compares several PDS sizes chosen for VC. ``$\sigma=2\%$'' 
% is defined in Table~\ref{tab:symbol}. 
% =======
Deduplication efficiency is defined as the percent of duplicate chunks which are removed
compared to a perfect scheme which detects and removes  all duplicates. 
% so only global perfect deduplication can have 100\% efficiency.
% We also compare several PDS sizes chosen for VC.
%(see Table~\ref{tab:symbol} for definition of ``$\sigma=2\%$'')
%we allocate the space of distributed shared memory  to accommodate 2\%
%of data chunks shared among VMs, as defined in 
Notice $\sigma$ is the percentage of unique chunks selected in PDS.
With $\sigma =2\%$, Column 3 shows its 
%memory usage for PDS index lookup per machine is about 100MB per machine and  the 
deduplication efficiency can 
reach over 96\%.
When $\sigma=4\%$, Column 4 shows that the deduplication efficiency can reach 96.6\% while 
more index space is allocated  per machine for  PDS fingerprint lookup.
% increases from 100MB to 200MB per physical machine as discussed later. 
%and  the deduplication efficiency can reach 96.489\%.
%When $\sigma=4\%$, the deduplication efficiency can reach 97.027\% while space consumption increases to 200MB per machine. 
The loss of efficiency in VC is caused by the restriction of the physical memory available
in the cluster for fast in-memory PDS index lookup. 
%In the extreme case if we put all the fingerprint into PDS 
%index ($\sigma=100\%$) then VC will have 100\% efficiency.
SRB in Column 5 can deliver up to 97.86\% deduplication efficiency, which is slightly better than VC.
Thus this represents a tradeoff that VC provides better fault tolerance and fast approximate deletion
with a slight reduction in deduplication efficiency.

% SRB can deliver up to 97.935\% deduplication efficiency which is slightly better than VC.
% Thus this represents a tradeoff: VC provides better fault tolerance and fast
% approximate deletion with competitive (but slightly lower) deduplication
% efficiency.
% >>>>>>> 4dcd68612b052c0f890ed30361a9ed03771b5df1
%its loss of efficiency is caused by the routing of data chunks which restricts the search scope
%of duplicate detection.
%SRB provides slightly better deduplication efficiency than our VC approach, which we accept
%because we are rewarded by architecture-wise benefits such as fault-tolerance and fast 
%Though not our primary goal, this result does show that VC can remove competitive amount of duplicates.

Column 2 of Table~\ref{tab:efficiency-comparison} shows deduplication efficiency of VC without 
using PDS and it achieves 93.02\% and  missed duplicates take  up-to  3.56\% of the original space, which
is about 356GB per physical machine. 
Column 1 of Table~\ref{tab:efficiency-comparison} shows deduplication efficiency of VC without 
local similarity search.
%There is a big efficiency drop in this curve when the number of VMs is about 30.
The reason similarity search provides such an improvement is that
%Without this (i.e. when we only 
there are VMs in which
data  segments are moved to another location on disk, for example when a file is rewritten
rather than modified in place,  
and a dirty-bit or offset-only based detection would not be able to detect such a movement.
We have found that in
approximately 1/3 of the VMs in our dataset this movement happens frequently.
In general, adding local similarity-guided search increases deduplication efficiency from 94\% to over 96\%.
That is one significant improvement compared to the work in ~\cite{WeiZhangIEEE}
which uses the parent segment at the same offset to detect duplicates instead of similarity-guided search.
% In general, our experiments show that
% dirty-bit detection at the segment level can reduce the data size to about 24.14\% of original data, 
% which leads about a 75.86\% reduction.
% Similarity-guided local search can further reduce the data size
% to about 12.05\% of original size, namely it delivers an additional 12.09\% reduction.
% The PDS-based global duplicate detection with $\sigma=2\% $
% can reduce the size further to 8.6\% of original size, namely it 
% delivers an additional 3.45\% reduction.

@@

@@

\subsection{Resource Usage and Processing Time}

\comments{
\begin{table}
    \begin{tabular}{|c|c|}
    \hline
    PDS replication degree & Disk usage per machine  \\ \hline
    3                      & 3065GB             \\ \hline
    4                      & 3072GB             \\ \hline
    5                      & 3079GB             \\ \hline
    6                      & 3086GB             \\ \hline
    7                      & 3093GB             \\ \hline
    \end{tabular}
\caption{Storage space cost per machine for DFS under different PDS replication degree}
\label{tab:replication_cost}
\end{table}
}

@@

{\bf Processing Time breakdown without I/O throttling}.
Table~\ref{tab:vc_srb_combined} shows
the  average  time breakdown for processing a dirty VM segment in milliseconds
under VC and Stateless Routing with Binning (SRB). 
VC uses $\sigma=2\%$ and $4\%$.
The overall processing latency of SRB is about 23.9\% slower than VC.
For VC, the change of $\sigma$ does not significantly affect the overall backup speed as
PDS lookup takes only a small amount of time. 
It has a breakdown of processing time.
``Read/Write'' includes snapshot reading and writing from disk, and 
updating of the metadata.
``Network'' includes the cost of transferring raw and meta data from one machine to another during 
snapshot read and write.
``Index Lookup'' is the disk, network and CPU time during fingerprint comparison.
This includes PDS data lookup for VC and index lookup from disk in SRB. 
The network transfer time for VC and SRB is about the same, because the 
amount of raw data they transfer is comparable.
SRB spends slightly more time for snapshot read/write because during each snapshot  backup, SRB involves many small bins,
 while VC only involves few containers with a bigger size. Thus, there are more opportunities for I/O aggregation in VC to reduce seek time.
SRB also has a higher cost for index access and fingerprint comparison because most chunk fingerprints are routed
to remote machines for comparison while   VC handles most chunk fingerprints locally.

@@

\comments{
 \begin{figure}
\vspace{-0.1in}
     \centering
     \includegraphics[width=3in]{figures/leakage}
     \caption{Accumulated storage leakage by approximate snapshot deletions ($\Delta u/u=0.025$)}
     \label{fig:leakage}
 \end{figure}
}
\begin{table}
\centering
\begin{small}
    \begin{tabular}{|c|c|c|c|c|c|}
    \hline
Del. step   &     1 &  3  & 5 & 7  & 9 \\ 
\hline   
     Estimated& .02\% & .06\% & .10\%& .14\% & .18\%\\ 
\hline   
     Measured & .01\% & .055\% & .09\% & .12\% & .15\% \\
\hline   
    \end{tabular}
\end{small}
     \caption{Accumulated storage leakage by approximate snapshot deletions ($\Delta u/u=0.025$)}
     \label{tab:leakage}
\end{table}

>>>>

\subsection{Deduplication Efficiency and Memory Usage}
\label{sect:evaldedup}

@@

\comments{
alg     efficiency
\scriptsize{}VC-2\%\tiny{}No\ Similarity        94.3132243792193
\scriptsize{}VC-NoPDS   93.0202119345744
\scriptsize{}VC-2\%PDS  96.0127647380588
\scriptsize{}VC-4\%PDS  96.5784837160482
\scriptsize{}SRB        97.8625610883421
}

\comments{
old table, delete 4%
    \begin{tabular}{|c|c|c|c|c|c|}
    \hline 
&
     \multicolumn{4}{|c|}{VC}  & VO \\ \hline
&
     2\%PDS  &  No PDS &2\%PDS     & 4\%PDS & SRB  \\ 
&     no-simi&   &&  &    \\ \hline
Dedup &     94.31\%& 93.02\%  & 96.01\% & 96.58\% & 97.86\% \\ \hline
Mem(MB) &     108&  46 & 118  & 190 & 2400 \\ \hline
    \end{tabular}

}
\begin{table}[htb]
\centering
\begin{small}
    \begin{tabular}{|c|c|c|c|c|}
    \hline 
&
     \multicolumn{3}{|c|}{VC}  & VO \\ \hline
&
     2\%No PDS &2\%PDS     & 2\%PDS & SRB  \\ 
&        &&     no-simi & \\ \hline
Dedup &      93.02\%  & 96.01\% & 94.31\% & 97.86\% \\ 
efficiency &          &       &           &      \\    \hline 
Memory(MB) &      46 & 118  & 108 & 2400 \\ \hline
    \end{tabular}
\end{small}
    \caption{Deduplication efficiency and per-machine memory usage for different VC settings and
        Stateless Routing with Binning (SRB).}
    \label{tab:efficiency-comparison}
\end{table}

Table~\ref{tab:efficiency-comparison} shows the deduplication efficiency and per-machine memory usage for SRB and VC with different 
settings.
% <<<<<<< HEAD
% Our deduplication efficiency is defined as the percent of duplicate chunks
% which are detected and deduplicated, so only global perfect deduplication can have 100\% efficiency.
% The figure also compares several PDS sizes chosen for VC. ``$\sigma=2\%$'' 
% is defined in Table~\ref{tab:symbol}. 
% =======
Deduplication efficiency is defined as the percent of duplicate chunks which are removed
compared to a perfect scheme which detects and removes  all duplicates. 
% so only global perfect deduplication can have 100\% efficiency.
% We also compare several PDS sizes chosen for VC.
%(see Table~\ref{tab:symbol} for definition of ``$\sigma=2\%$'')
%we allocate the space of distributed shared memory  to accommodate 2\%
%of data chunks shared among VMs, as defined in 
Notice $\sigma$ is the percentage of unique chunks selected in PDS.
With $\sigma =2\%$, Column 3 shows its 
%memory usage for PDS index lookup per machine is about 100MB per machine and  the 
deduplication efficiency can 
reach over 96\%. 
%When $\sigma=4\%$, Column 4 shows that the deduplication efficiency can reach 96.6\% while 
%more index space is allocated  per machine for  PDS fingerprint lookup.
The loss of efficiency in VC is caused by the restriction of the physical memory available
in the cluster for fast in-memory PDS index lookup. 
Memory usage per machine is low because each machine only hosts $1/p$  of  index for PDS plus some buffer space. 
%In the extreme case if we put all the fingerprint into PDS 
%index ($\sigma=100\%$) then VC will have 100\% efficiency.
SRB in Column 5 can deliver up to 97.86\% deduplication efficiency, which is slightly better than VC.
Thus this represents a tradeoff that VC provides better fault tolerance and fast approximate deletion
with a slight reduction in deduplication efficiency.
Memory usage per machine in SRB includes the Bloom filter space to access the on-disk index
and  cache for frequently or recently accessed chunk index.  
% SRB can deliver up to 97.935\% deduplication efficiency which is slightly better than VC.
% Thus this represents a tradeoff: VC provides better fault tolerance and fast
% approximate deletion with competitive (but slightly lower) deduplication
% efficiency.
% >>>>>>> 4dcd68612b052c0f890ed30361a9ed03771b5df1
%its loss of efficiency is caused by the routing of data chunks which restricts the search scope
%of duplicate detection.
%SRB provides slightly better deduplication efficiency than our VC approach, which we accept
%because we are rewarded by architecture-wise benefits such as fault-tolerance and fast 
%Though not our primary goal, this result does show that VC can remove competitive amount of duplicates.

Column 2 of Table~\ref{tab:efficiency-comparison} shows deduplication efficiency of VC without 
using PDS and it achieves 93.02\% and  missed duplicates take  up-to  3.56\% of the original space, which
is about 356GB per physical machine. 
Column 4 of Table~\ref{tab:efficiency-comparison} shows deduplication efficiency of VC which only uses
the same offset in the parent segment to look for duplicates  but does not use 
the fingerprint-guided similarity search. 
%There is a big efficiency drop in this curve when the number of VMs is about 30.
The reason similarity search provides such an improvement is that
%Without this (i.e. when we only 
there are VMs in which
data  segments are moved to another location on disk, for example when a file is rewritten
rather than modified in place,  
and a dirty-bit or offset-only based detection would not be able to detect such a movement.
We have found that in
approximately 1/3 of the VMs in our dataset this movement happens frequently.
In general, adding local similarity-guided search increases deduplication efficiency from 94\% to over 96\%.
That is one significant improvement compared to the work in ~\cite{WeiZhangIEEE}
which uses the parent segment at the same offset to detect duplicates instead of similarity-guided search.
% In general, our experiments show that
% dirty-bit detection at the segment level can reduce the data size to about 24.14\% of original data, 
% which leads about a 75.86\% reduction.
% Similarity-guided local search can further reduce the data size
% to about 12.05\% of original size, namely it delivers an additional 12.09\% reduction.
% The PDS-based global duplicate detection with $\sigma=2\% $
% can reduce the size further to 8.6\% of original size, namely it 
% delivers an additional 3.45\% reduction.

@@

The overall memory usage of VC for each physical machine is very small, which is not visible to other hosted cloud services.
The 2.4GB memory usage of SRB is very visible to other services even each physical machine has 32GB memory.

@@

\subsection{More on Resource Usage and Processing Time}

We show  more experimental results on resource usage and processing time with SRB.
\comments{
\begin{table}
    \begin{tabular}{|c|c|}
    \hline
    PDS replication degree & Disk usage per machine  \\ \hline
    3                      & 3065GB             \\ \hline
    4                      & 3072GB             \\ \hline
    5                      & 3079GB             \\ \hline
    6                      & 3086GB             \\ \hline
    7                      & 3093GB             \\ \hline
    \end{tabular}
\caption{Storage space cost per machine for DFS under different PDS replication degree}
\label{tab:replication_cost}
\end{table}
}

@@

{\bf Processing Time breakdown without I/O throttling}.
Table~\ref{tab:vc_srb_combined} shows
the  average  time breakdown for processing a dirty VM segment in milliseconds
under VC and Stateless Routing with Binning (SRB). 
%VC uses $\sigma=2\%$ and $4\%$.
VC uses $\sigma=2\%$. 
The overall processing latency of SRB is about 23.9\% slower than VC.
For VC, the change of $\sigma$ does not significantly affect the overall backup speed as
PDS lookup takes only a small amount of time. 
It has a breakdown of processing time.
``Read/Write'' includes snapshot reading and writing from disk, and 
updating of the metadata.
``Network'' includes the cost of transferring raw and meta data from one machine to another during 
snapshot read and write.
``Index Lookup'' is the disk, network and CPU time during fingerprint comparison.
This includes PDS data lookup for VC and index lookup from disk in SRB. 
The network transfer time for VC and SRB is about the same, because the 
amount of raw data they transfer is comparable.
SRB spends slightly more time for snapshot read/write because during each snapshot  backup, SRB involves many small bins,
 while VC only involves few containers with a bigger size. Thus, there are more opportunities for I/O aggregation in VC to reduce seek time.
SRB also has a higher cost for index access and fingerprint comparison because most chunk fingerprints are routed
to remote machines for comparison while   VC handles most chunk fingerprints locally.

@@

\input{paper_eval_del}
\comments{
 \begin{figure}
\vspace{-0.1in}
     \centering
     \includegraphics[width=3in]{figures/leakage}
     \caption{Accumulated storage leakage by approximate snapshot deletions ($\Delta u/u=0.025$)}
     \label{fig:leakage}
 \end{figure}
}
\begin{table}
\centering
\begin{small}
    \begin{tabular}{|c|c|c|c|c|c|}
    \hline
Del. step   &     1 &  3  & 5 & 7  & 9 \\ 
\hline   
     Estimated& .02\% & .06\% & .10\%& .14\% & .18\%\\ 
\hline   
     Measured & .01\% & .055\% & .09\% & .12\% & .15\% \\
\hline   
    \end{tabular}
\end{small}
     \caption{Accumulated storage leakage by approximate snapshot deletions ($\Delta u/u=0.025$)}
     \label{tab:leakage}
\end{table}
